---
title: "HW6"
author: "Ben Rochford"
format: html
editor: visual
---

## HW6

Setup

```{r}
library(tidyverse)
library(broom)

theme_set(theme_linedraw(base_family = "Avenir Next Condensed"))
```

## 6.2 Balance and Overlap

### 6.2.1

```{r}
d <- haven::read_dta("cattaneo2.dta")

d <- d |>  
  haven::zap_labels() |>             
  select(bweight, lbweight, mbsmoke, mmarried, mage, medu, fbaby, alcohol, mrace, nprenatal)

glimpse(d)
```

Finding difference in means of covariates for mbsmoke

```{r}
standard_diff_smoke_to_cov <- function(cov) {
  return(
    (mean(d[[cov]][d$mbsmoke == 1]) - mean(d[[cov]][d$mbsmoke == 0])) / sd(d[[cov]][d$mbsmoke == 1])
  )
}


cols_cov <- colnames(d)[colnames(d) != "mbsmoke"]
d_cov <- tibble(cov = cols_cov,
                diff = sapply(cols_cov, standard_diff_smoke_to_cov)) |> arrange(desc(abs(diff)))

d_cov
```

```{r}
ggplot(data = d_cov, mapping = aes(x = diff, y = reorder(cov, abs(diff)))) + 
  geom_segment(xend = 0, yend = d_cov$cov) + 
  geom_point() +
  geom_vline(xintercept = 0, linetype = "dashed")
```

The worst offenders, in terms of bias as potential confounders, are those towards the top of the chart and table.

## 6.3 Matching

### 6.3.1

You want to know whether practicing cursive improves your penmanship (on a 1-10 scale). You find that, among people who don’t practice cursive, average penmanship is 5, 10 people are left-handed, 2 are ambidextrous, and 88 are right-handed. Among people who do practice cursive, 6 are left-handed with average penmanship 7, 4 are ambidextrous with average penmanship 4, and 90 are right-handed with average penmanship 6.

a.  You want to create a set of weights that will *make the treated group match the control group on handedness*. Follow the process in [section 14.2](https://theeffectbook.net/ch-Matching.html#weighted-averages), paying attention to *why* certain numbers are going in certain positions. What weights will be given to the left, ambidextrous, and right-handed people *in the control group*?

    > In the control group, all will be weighted with 1.

b.  What weights will be given to the left, ambidextrous, and right-handed people *in the treated group*?

    ```{r}
    10 / 6  # 10 in untreated, 6 in treated
    2 / 4   # 2 in untreated, 4 in treated
    88 / 90 # 88 in untreated, 90 in treated
    ```

    > 1.667 for left, .5 for ambidextrous, 0.978 for right

c.  Use the weights from part b to calculate the *proportion of left-handed people in the treated group*, as well as the proportion of ambidextrous people and the proportion of right-handed people. If you don’t get 10%, 2%, and 88% (or very close with some rounding error), your weights are wrong, try again.

    ```{r}
    1.667 * (6/100)
    0.5 * (4/100)
    0.978 * (90/100)
    ```

d.  What is the weighted average penmanship score in the treated group?

    ```{r}
    ( (1.667*6*7) + (0.5*4*4) + (0.978*90*6) ) / 100
    ```

e.  What is the effect of practicing cursive that we would estimate using this data?

    ```{r}
    6.06134 - 5
    ```
