---
title: "HW1"
author: "Ben Rochford"
format: html
editor: visual
---

## 1.1

Estimand

OLS

Unit-specific quantity

Potential outcomes

Effect heterogeneity

### Setup (1.2)

```{r}
library(tidyverse)
library(gt)

d <- data.frame(
  T = c(0, 0, 1, 0, 0, 1, 1, 1),
  Y0 = c(5, 8, 5, 12, 4, 8, 4, 9),
  Y1 = c(5, 10, 3, 13, 2, 9, 1, 13), 
  id = LETTERS[1:8]
)

gt(d, rowname_col = "id") # gt is used for fancy printing of tables
```

## 1.3

Add outcomes and calculate the Average Treatment Effect

$Y = TY^1 + (1-T)Y^0$

```{r}
d <- d |> 
  mutate(Y = T * Y1 + (1 - T) * Y0)

gt(d, rowname_col = "id")
```

Now the ATE

```{r}
mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])
```

## 1.4

re sample T 3 times and run

```{r}
d <- d |> 
  mutate(T = sample(d$T)) |> 
  mutate(Y = T * Y1 + (1 - T) * Y0)

gt(d, rowname_col = "id")
mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])


d <- d |> 
  mutate(T = sample(d$T)) |> 
  mutate(Y = T * Y1 + (1 - T) * Y0)

gt(d, rowname_col = "id")
mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])


d <- d |> 
  mutate(T = sample(d$T)) |> 
  mutate(Y = T * Y1 + (1 - T) * Y0)

gt(d, rowname_col = "id")
mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])
```

The estimates are generally not very close to the "real" ATE

## 1.5

Find the needed sample size for statistical power

We know our ATE is .125, and we will assume reasonable values for other parameters

```{r}
power.t.test(delta=0.125, sd=1, sig.level=0.05, power=0.80, type="two.sample", alternative="two.sided")
```

Looks like the sample size needs to be a little over a thousand here, to have a statistical power of .8
