---
title: "HW10"
author: "Ben Rochford"
format: html
editor: visual
---

# HW10

Setup

```{r}
library(tidyverse)
library(panelr)
library(gapminder)
library(ggeffects)
```

### 10.1 - Reshaping data

```{r}
gap_wide <- gapminder::gapminder |> 
  select(continent, country, year, lifeExp, gdpPercap, pop) |> 
  pivot_wider(
    names_from = year, 
    values_from = c(lifeExp, gdpPercap, pop), 
    names_sep = ""
  )

gap_wide
```

Use the `panelr` package to turn the dataset back to its original long form and save it as `gap_long`

```{r}
gap_long <- long_panel(gap_wide, periods = 
                         seq(1952, 2007, by = 5)) |> 
  select(wave, continent, country, lifeExp, gdpPercap, pop)

gap_long
```

Use the `panelr` package to turn the `gap_long` into `gap_wide`, replicating `pivot_wider()`

```{r}
gap_wide <- widen_panel(gap_long)
gap_wide
```

### 10.2 - `lifeExp` visualization

```{r}
r10_countries <- gap_wide |> 
  sample_n(10) |> 
  long_panel(prefix = "_", periods = seq(1952, 2007, by = 5))

lifeYear <- ggplot(r10_countries, aes(x = wave, y = lifeExp, color = country)) +
  geom_line() + theme_light()

lifeYear
```

### 10.3 - Intra-class Correlation

What is the ICC for `lifeExp`, `pop`, and `gdpPercap` in the `gapminder` dataset?

```{r}
m_lifeExp <- lmer(lifeExp ~ (1 | country), data = gapminder, REML = FALSE)
summary(m_lifeExp)
```

`lifeExp`:

```{r}
119.40 / (119.40 + 47.36)
```

```{r}
m_pop <- lmer(pop ~ (1 | country), data = gapminder, REML = FALSE)
summary(m_pop)
```

`pop`:

```{r}
1.021e+16 / (1.021e+16  + 1.055e+15)
```

```{r}
m_gdpPercap <- lmer(gdpPercap ~ (1 | country), data = gapminder, REML = FALSE)
summary(m_gdpPercap)
```

`gdpPercap`:

```{r}
69177620 / (69177620 + 27934767)
```

### 10.4 - `WageData` `lmer`s

```{r}
data("WageData", package = "panelr")

WageData <- WageData |> 
  mutate(
    college = if_else(ed >= 16, 1L, 0L),  # college variable
    t0 = t - 1                             # start time at 0
  )
```

Log wage as a function of college and linear time

```{r}
lw_col_time = lmer(lwage ~ college + t + (1 | id), 
                   data = WageData, REML = FALSE)
summary(lw_col_time)
```

... plus a random slope on time

```{r}
lw_rt = lmer(lwage ~ college + t + (1 + t | id), 
             data = WageData, REML = FALSE)
summary(lw_rt)
```

... now with time as a quadratic

```{r}
lw_qt = lmer(lwage ~ college + t + I(t^2) + (1 + t + I(t^2) | id), 
             data = WageData, REML = FALSE)
summary(lw_qt)
```

**Comparing models with BIC**

```{r}
c(BIC(lw_col_time), BIC(lw_rt), BIC(lw_qt))
```

The lowest BIC score belongs to model 3 with the quadratic term - slightly beating the model with random slope on time included, where both of these models significantly beat model 1.

Report the estimated effect of college on log wages given the data and model. You can get this using `tidy()`, `summary()`, or any other function you prefer.

> According to model3, `college` increases log earnings by 0.312

Use `ggpredict() |> plot()` (or another approach) to plot predictions for a sample of 9 individuals over time.

```{r}
wageSample <- WageData |> 
  mutate(p = predict(lw_qt)) |> 
  group_by(id) |> nest() |> ungroup() |> 
  slice_sample(n = 9) |> 
  unnest(data)

ggplot(wageSample, aes(x = t0, y = p, color = factor(id))) +
         geom_line()
```
